---
title: "EstimateGenomeSize"
author: "Jessie Pelosi"
date: "2024-08-09"
output: html_document
---

This script was used to estimate the approximate haploid genome size of ≤i>Lygodium microphyllum</i> using short-read data. 

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

This script was used to estimate the approximate haploid genome size of ≤i>Lygodium microphyllum</i> using short-read data. We will use k-mer frequencies counted with Jellyfish. 

This block uses all short-read data, which is approximately 60x coverage. We will use multiple different kmers (21, 31, 41, 51, and 61). 

```{r}
###############All short read data ######################

# 21 mers
spec1_21 <- read.table("./shortreads/21mer_reads.histo")
plot(spec1_21[3:200,], type="l", xlab = "Coverage", ylab = "Frequency", main = "Jellyfish 21mers")
png(file="21mer_jellyfish_shortreads.png", width = 900, height =750)
dev.off()
points(spec1_21[3:200,])
sum(as.numeric(spec1_21[3:200,1]*spec1_21[3:200,2]))
max(spec1_21[10:10000,2]) #52x 
Gsize21 <- sum(as.numeric(spec1_21[1:10000,1]*spec1_21[1:10000,2]))/52
# 21 mer : 3.58 Gb @ 52x
#(sum(as.numeric(spec1_21[2:75,1]*spec1_21[2:75,2]))) / (sum(as.numeric(spec1_21[2:10000,1]*spec1_21[2:10000,2])))

# 31 mers
spec1_31 <- read.table("./shortreads/31mer_reads.histo")
plot(spec1_31[3:200,], type="l", xlab = "Coverage", ylab = "Frequency", main = "Jellyfish 31mers")
png(file="31mer_jellyfish_shortreads.png", width = 900, height =750)
dev.off()
points(spec1_31[3:200,])
sum(as.numeric(spec1_31[3:200,1]*spec1_31[3:200,2]))
max(spec1_31[10:10000,2]) #47x
Gsize31 <- sum(as.numeric(spec1_31[1:10000,1]*spec1_31[1:10000,2]))/47
# 31mer : 4.1 Gb @ 47x

# 41 mers
spec1_41 <- read.table("./shortreads/41mer_reads.histo")
plot(spec1_41[3:200,], type="l", xlab = "Coverage", ylab = "Frequency", main = "Jellyfish 41mers")
png(file="41mer_jellyfish_shortreads.png", width = 900, height =750)
dev.off()
points(spec1_41[3:200,])
sum(as.numeric(spec1_41[3:200,1]*spec1_41[3:200,2]))
max(spec1_41[10:10000,2]) #42x
Gsize41 <- sum(as.numeric(spec1_41[1:10000,1]*spec1_41[1:10000,2]))/42
# 41mer : 4.4 Gb @ 42x 

#51 mers
spec1_51 <- read.table("./shortreads/51mer_reads.histo")
plot(spec1_51[3:200,], type="l", xlab = "Coverage", ylab = "Frequency", main = "Jellyfish 51mers")
png(file="51mer_jellyfish_shortreads.png", width = 900, height =750)
dev.off()
points(spec1_51[3:200,])
sum(as.numeric(spec1_51[3:200,1]*spec1_51[3:200,2]))
max(spec1_51[4:10000,2]) #37x
Gsize51 <- sum(as.numeric(spec1_51[1:10000,1]*spec1_51[1:10000,2]))/37
# 51mer : 4.6 Gb @ 37x

#61 mer
spec1_61 <- read.table("./shortreads/61mer_reads.histo")
plot(spec1_61[3:200,], type="l", xlab = "Coverage", ylab = "Frequency", main = "Jellyfish 61mers")
png(file="61mer_jellyfish_shortreads.png", width = 900, height =750)
dev.off()
points(spec1_61[3:200,])
sum(as.numeric(spec1_61[3:200,1]*spec1_61[3:200,2]))
max(spec1_61[4:10000,2]) #32x
Gsize61 <- sum(as.numeric(spec1_61[1:10000,1]*spec1_61[1:10000,2]))/32
# 61mer : 4.8 Gb @ 32x

```

To summarize: the approximate haploid genome sizes for each of these kmers is as follows 
```{r}
kmers <- c("21", "31", "41","51", "61")
sizes <- c(Gsize21, Gsize31, Gsize41, Gsize51, Gsize61)
sum <- data.frame(kmers,sizes)
colnames(sum) <- c("kmer", "Haploid Genome Size Estimate")
kable(sum, align = "c") %>% 
  kable_styling(position = "center")

```
